---
title: "Heart Disease prediction"
output:
  html_document:
    df_print: paged
---

```{r}
library(neuralnet)
library(utils)
library(psych)
library(ggplot2)
library(corrplot)
library(neuralnet)
library(randomForest)
library(gbm)
library(e1071)
library(pROC)
```

```{r}
heart_data <-read.csv("C:/Users/testuser/Downloads/heart.csv")
```

```{r}
summary(heart_data)
```
```{r}
describe(heart_data)
```

```{r}
corrplot::cor.mtest(heart_data)
```

```{r}
#checking the severity of chest pain in male and female
ggplot(heart_data,aes(cp,col='red'))+geom_histogram(bins=10)+facet_wrap(vars(output))

```

```{r}
# For example, visualizing 'age' distribution 
ggplot(heart_data, aes(x = factor(age), fill = factor(age))) + 
  geom_bar() +  
  labs(title = "Age Distribution", x = "Age", y = "Count") +  
  theme_minimal() 
```

#The likelihood of having a heart attack among both populations in a scale from 0-1 the closer to 1 the higher the probability of a heart event
```{r}
ggplot(data = heart_data, aes(x = factor(sex), fill = factor(sex))) +
  geom_bar() +
  scale_fill_manual(values = c("blue", "pink"), name = "Gender", labels = c("Female", "Male")) + 
  labs(title = "Distribution of Males and Females", x = "Gender", y = "Count") +
  theme_minimal() +
  labs(x = "Gender", y = "Count") +  
  theme(legend.position = "top") 
```



```{r}
# Chest Pain distribution 
ggplot(heart_data, aes(x = factor(cp), fill = factor(cp))) + 
  geom_bar() +
  labs(title = "Chest Pain Distribution", x = "CP", y = "Count") +
  theme_minimal()
```

#The graphic depicted above shows the ages of both men and women where there is more propensity and higher frequencies of a heart event for men between 55 – 65 and for women the likelihood increases after ag 65.
```{r}
# Assuming 0 represents females and 1 represents males
ggplot(data = heart_data, aes(x = age, fill = factor(sex))) +
  geom_histogram(position = "stack", color = "black", bins = 30) +
  scale_fill_manual(values = c("blue", "pink"), name = "Gender", labels = c("Female", "Male")) +  
  labs(title = "Age Distribution by Gender", x = "Age", y = "Count") +
  theme_minimal() +
  theme(legend.position = "top")  
```


```{r}
# Correlation plot
corhd <- cor(heart_data[, 1:13])
corrplot(corhd, type = "upper", order = "hclust", tl.cex = 0.7)
```

```{r}
# Blood Pressure and cholestrol 
x1 <- ggplot(data = heart_data, aes(x = chol, y = trtbps, col = factor(sex))) +
  geom_point() + geom_smooth(method = "lm", se = FALSE)
x1
```



#the heart dataset for predicting heart failure based on several input variables like age, sex, chest pain type (cp), resting blood pressure (trtbps), cholesterol levels (chol), fasting blood sugar (fbs), resting electrocardiographic results (restecg), maximum heart rate achieved (thalachh), exercise-induced angina (exng), ST depression induced by exercise relative to rest (oldpeak), slope of the peak exercise ST segment (slp), number of major vessels (caa) colored by fluoroscopy, and thalassemia (thall).
```{r}
# Assuming 'heart_data' contains predictors and 'output' is the target variable
# Split the data into training and testing sets
set.seed(123)
train_indices <- sample(nrow(heart_data), 0.7 * nrow(heart_data))
train_data <- heart_data[train_indices, ]
test_data <- heart_data[-train_indices, ]
```

```{r}
# Logistic Regression
# Fit the logistic regression model
model_logistic <- glm(output ~ ., data = train_data, family = "binomial")

# Random Forest
# Fit the random forest model
model_rf <- randomForest(output ~ ., data = train_data, ntree = 100)

# Gradient Boosting
# Fit the gradient boosting model
model_gbm <- gbm(output ~ ., data = train_data, n.trees = 100, distribution = "bernoulli")

# SVM
# Fit the SVM model
model_svm <- svm(output ~ ., data = train_data, kernel = "radial")
```

```{r}
# Predictions
# Predict on test set
pred_logistic <- predict(model_logistic, newdata = test_data, type = "response")
pred_rf <- predict(model_rf, newdata = test_data, type = "response")
pred_gbm <- predict(model_gbm, newdata = test_data, n.trees = 100)
pred_svm <- predict(model_svm, newdata = test_data)

# Plots (ROC curves for evaluation)
roc_logistic <- roc(test_data$output, pred_logistic)
roc_rf <- roc(test_data$output, pred_rf)
roc_gbm <- roc(test_data$output, pred_gbm)
roc_svm <- roc(test_data$output, pred_svm)

# Plotting ROC curves
plot(roc_logistic, col = "blue", main = "ROC Curves")
plot(roc_rf, col = "red", add = TRUE)
plot(roc_gbm, col = "green", add = TRUE)
plot(roc_svm, col = "orange", add = TRUE)
legend("bottomright", legend = c("Logistic Regression", "Random Forest", "Gradient Boosting", "SVM"),
       col = c("blue", "red", "green", "orange"), lty = 1)
```
```{r}
# Assuming 'test_data$output' contains the actual output values and 'pred_logistic'/'pred_rf'/'pred_gbm'/'pred_svm' contain predicted values for Logistic Regression, Random Forest, Gradient Boosting, and SVM respectively

# Confusion matrix for Logistic Regression
conf_matrix_logistic <- table(test_data$output, ifelse(pred_logistic > 0.5, 1, 0))
accuracy_logistic <- sum(diag(conf_matrix_logistic)) / sum(conf_matrix_logistic)

# Confusion matrix for Random Forest
conf_matrix_rf <- table(test_data$output, ifelse(pred_rf > 0.5, 1, 0))
accuracy_rf <- sum(diag(conf_matrix_rf)) / sum(conf_matrix_rf)

# Confusion matrix for Gradient Boosting
conf_matrix_gbm <- table(test_data$output, ifelse(pred_gbm > 0.5, 1, 0))
accuracy_gbm <- sum(diag(conf_matrix_gbm)) / sum(conf_matrix_gbm)

# Confusion matrix for SVM
conf_matrix_svm <- table(test_data$output, ifelse(pred_svm > 0.5, 1, 0))
accuracy_svm <- sum(diag(conf_matrix_svm)) / sum(conf_matrix_svm)

# Printing accuracies
cat("Logistic Regression Accuracy:", accuracy_logistic, "\n")
cat("Random Forest Accuracy:", accuracy_rf, "\n")
cat("Gradient Boosting Accuracy:", accuracy_gbm, "\n")
cat("SVM Accuracy:", accuracy_svm, "\n")

```








